% A simple template for LaTeX documents
% 
% To produce pdf run:
%   $ pdflatex paper.tex 
%

\documentclass[12pt]{article}

% Change margin size
\usepackage[margin=1in]{geometry}   


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\section{Literature Review}


\subsection{SQL query optimization}

SQL is nice because the unit of analysis is a single query. Thus the query
optimizer only has to figure out how to do this one query efficiently,
although nested subqueries could add complexity.

\cite{Chaudhuri:1998:OQO:275487.275492} gives an overview of query
optimization for relational databases. The execution plan is a description
of the physical operators, ie. table scan, that takes one or more input
data streams and outputs another stream.

The author distinguishes Query optimization from execution. The former
decides the execution plan, the latter executes it.

Query optimization needs: a \textbf{search space} of possible solutions,
\textbf{cost estimation} for each solution, and an \textbf{enumeration
algorithm} to explore the different solutions.

System R was an early database technology that introduced some of the
query optimization techniques.

\subsubsection{relation to my work}

An R expression, particularly something like tables piped through dplyr, is
like a query. If we have an expensive vectorized function \texttt{f} and a
large data structure \texttt{x} then \texttt{f(head(x))} is semantically
equivalent to, but much cheaper than, \texttt{head(f(x))}. The reason is
because the filtering happens first.

The example above suggests that we rewrite code such that all filtering
happens as early as possible. We could also rewrite iterative filters in a
way that removes the intermediate data structures, but I'm not sure that
comes up enough to justify it. Then we need to look at some corpus of R
code to see how people actually use it, instead of my ad hoc experience of
how people use it.

R differs from SQL because it is executable code rather than a declarative
language. My work centers on modifying code for performance while
preserving the semantics. What is the search space?  We can change
\texttt{lapply} to \texttt{mclapply}. We can evaluate expressions
asynchronously. We can use a database for the \texttt{split, apply,
combine} paradigm as described in \cite{wickham2011split}. This paradigm is
actually quite amenable to different computational approaches- the
semantics are clear.

\end{document}
